The \dealii Library, Version 9.0

### 2018-04-24 9.0/paper.tex
Small changes here and there.

\usepackage{fontenc}
\usepackage{graphicx}

@@

\newcommand{\specialword}[1]{\texttt{#1}}
\newcommand{\dealii}{{\specialword{deal.II}}}
\newcommand{\pfrst}{{\specialword{p4est}}}
\newcommand{\trilinos}{{\specialword{Trilinos}}}
\newcommand{\aspect}{\specialword{Aspect}}
\newcommand{\petsc}{\specialword{PETSc}}
\newcommand{\cmake}{{\specialword{CMake}}}
\newcommand{\autoconf}{{\specialword{autoconf}}}

@@

\title{The \dealii{} Library, Version 9.0}

@@

\begin{abstract}
  This paper provides an overview of the new features of the finite element
  library \dealii{} version 9.0.
\end{abstract}

@@

\dealii{} version 9.0.0 was released April X, 2018.
\marginpar{Update the date}
This paper provides an
overview of the new features of this major release and serves as a citable
reference for the \dealii{} software library version 9.0. \dealii{} is an
object-oriented finite element library used around the world in the
development of finite element solvers. It is available for free under the
GNU Lesser General Public License (LGPL) from the \dealii{} homepage at
\url{http://www.dealii.org/}.

The major changes of this release are:
\begin{itemize}
\item xxxx something with manifolds xxx
\item Support for particle-in-cell methods
\item Dedicated support for automatic and symbolic differentiation.
\item Interfaces to more external libraries and programs.
\item C++11 is now both required and used.
\item Support for GPU computations.
\end{itemize}
These will all be discussed in more detail in the
following section. In addition, this release contains the following changes:
\begin{itemize}
\item \dealii{} has made extensive use of both the Clang-Tidy \cite{clang-tidy}
  and Coverity Scan \cite{coverity} static analysis tools for detecting bugs
  and other issues in the code. For example, around 260 issues were detected and
  fixed using the latter tool.
      %% > Is there somewhere we can cite for either of these tools? It would
      %% > be nice to recognize the department of homeland security for offering
      %% The contract with homeland security was completed in 2009.

@@

\item A new python tutorial program tutorial-1; as well as
  updates to step-37. In addition, the separate code
  gallery of \dealii{} has gained a number of new entries.
\end{itemize}
Beyond these changes, the changelog lists more than 330 other features and bugfixes.

@@

This release of \dealii{} contains a number of large and significant changes
that will be discussed in the following sections. It of course also contains a
vast number of smaller changes and added functionality; the details of these
can be found
\href{https://www.dealii.org/developer/doxygen/deal.II/changes_between_8_5_and_9_0.html}{
in the file that lists all changes for this release}, see \cite{changes90}.
(The file is also linked to from the web site of each release as well as
the release announcement.)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{xxx something with Manifolds xxx}
\label{sec:manifolds}

@@

(i) separation between manifold and boundary ids, (ii) manifold
smoothing via the transfinite mapping, (iii) maybe: grid generator functions already attach manifolds?

Previous text:

Manifold smoothing: The manifold smoothing algorithms applied in the
Triangulation class and MappingQGeneric have been changed from the old
Laplace-style smoothing to a transfinite interpolation that linearly
blends between the descriptions on the faces around a cell. The old
transformation introduced boundary layers inside cells that prevented
convergence rates from exceeding \(3.5\) in the global \(L^2\) errors on typical
settings. This change also considerably improves mesh quality on settings
where curved descriptions are only applied to the boundary rather than
the whole volume.

@@

\marginpar{Rene: Please edit as appropriate}
While \dealii{} is a package intended to solve problems with the
finite element method -- i.e., using continuous or discontinuous
\textit{fields} --, is often convenient in fluid dynamics problems to
couple the continuum description of phenomena with particles. These
particles, advected along with the numerical approximation of the flow
field, are then either used to visualize properties of the flow, or to
advect material properties such as the viscosity of inhomogenous
mixtures of fluids. If each particle is associated with the cells
of a mesh, these methods are often referred to as particle-in-cell (PIC).

\dealii{} now has a dedicated particles module. The module provides a base class
\texttt{Particle} that represents a particle with position, an ID number and
a variable number of properties. They are jointly represented by a \texttt{ParticleHandler}
class that manages the storage and handling of all particles. In
parallel simulations, this class also
distributes the particles among the subdomains of the parallel process
and supports efficient data transfer during mesh refinement and
checkpoint/restart phases.

A much more detailed view of the underlying algorithms can be found in
\cite{GLHPB18}. A longer report is at \cite{GHPB16}. The
implementation here originated in the \aspect{} code, see \cite{KHB12,HDGB17}.

@@

\dealii{} has had a tutorial program (step-33) since 2007 that
demonstrates this technique based on the Trilinos Sacado package, but
the functionality was not available pervasively throughout
deal.II. This has changed with release 9.0 where support for
differentiation has been collected in one module that provides both
automatic and symbolic differentiation. Currently, this module relies
on either the ADOL-C and Sacado libraries, through a unified
interface. In practice, this support means that ADOL-C and Sacado data
types can be used in the \texttt{FEValues}, \texttt{FEValuesViews},
and related classes that are generally used to assemble linear systems
and right hand sides.

@@

\dealii{} has always tried to leverage high-quality implementations of
algorithms available through other open source software, rather than
re-implementing their functionality. (A list of interfaces to other
packages is given in Section~\ref{sec:cite}.) As part of the current
release, we have written several new interfaces as discussed in the following.

@@

  \dealii{} now has interfaces to SUNDIALS's ARKode, IDA, and KINSOL sub-packages.
  ARKode is a solver library that provides adaptive-step time
  integration. IDA is a package for the solution of differential-algebraic
  equations systems in the form $F(t,y,y')=0$. KINSOL is solver for nonlinear
  algebraic systems.

@@

\paragraph*{nanoflann, a library for building and querying
  $k$-d trees of datasets.} Operations such as finding the vertex
or cell closest
to a given evaluation point occur frequently in many applications that
use unstructured meshes. While the naive algorithm is linear in the
number of vertices or cells, many such operations can be made
significantly faster by building a $k$-d tree data structure that
recursively subdivides a $k$ dimensional space. The nanoflann library
provides such a data structure and allows querying it, either for
closest points (e.g., when finding the closest vertex) or for
searching the points that fall within a radius of a target point. This
functionality is now available via \dealii{} interfaces.

\paragraph*{Assimp, the Open Asset Import Library.}
  Assimp can be used to read about 40 different 3D graphics formats. A
  subset of these formats can be now be read from within \dealii{} to
  generate two-dimensional meshes, possibly embedded in a
  three-dimensional space.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \subsection{Use of C++11}
  \label{sec:cxx11}
\dealii{} first offered support for a subset of C++11 features in
version 6.2, released in 2009. The current release is the first to
\emph{require} a C++11 compiler.

Many parts of the code base have been rewritten to both support and
use the new features of C++11. In particular, \dealii{} now makes
extensive use of move semantics as well as range-based \texttt{for}
loops with \texttt{auto} type deduction of iterator variables. We have
also largely replaced \texttt{push\_back()} by
\texttt{emplace\_back()} when adding elements to collections more
efficiently.

@@

@@

\dealii{} has a separate ``code gallery'' that
consists of programs shared by users as examples of what can be
done with \dealii{}. While not part of the release process, it is nonetheless
worth mentioning that the set of new programs since the last release covers
the following topics:
\begin{itemize}
\item The multipoint flux mixed finite element method (MFMFE) applied
  to the Darcy problem of porous media flow;
\item A linearized active skeletal muscle model with application to
  the simulation concentric contraction of the human biceps brachii;
\item A parallel implementation of the Local Discontinuous Galerkin
  (LDG) method applied to the Poisson equation.
\end{itemize}
With these additions, the code gallery now contains 10 different applications.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{How to cite \dealii{}}\label{sec:cite}

In order to justify the work the developers of \dealii{} put into this
software, we ask that papers using the library reference one of the
\dealii{} papers. This helps us justify the effort we put into it.

There are various ways to reference \dealii{}. To acknowledge the use of
the current version of the library, \textbf{please reference the present
document}. For up to date information and bibtex snippets for this document
see:
\begin{center}
 \url{https://www.dealii.org/publications.html}
\end{center}

The original \texttt{\dealii{}} paper containing an overview of its
architecture is \cite{BangerthHartmannKanschat2007}. If you rely on
specific features of the library, please consider citing any of the
following:
\begin{itemize}
 \item For geometric multigrid: \cite{Kanschat2004,JanssenKanschat2011};
 \item For distributed parallel computing: \cite{BangerthBursteddeHeisterKronbichler11};
 \item For $hp$ adaptivity: \cite{BangerthKayserHerold2007};
  \item For partition-of-unity (PUM) and enrichment methods of the
    finite element space: \cite{Davydov2016};
 \item For matrix-free and fast assembly techniques:
   \cite{KronbichlerKormann2012};
 \item For computations on lower-dimensional manifolds:
   \cite{DeSimoneHeltaiManigrasso2009};
 \item For integration with CAD files and tools:
   \cite{HeltaiMola2015};
 \item For \texttt{LinearOperator} and \texttt{PackagedOperation} facilities:
   \cite{MaierBardelloniHeltai-2016-a,MaierBardelloniHeltai-2016-b}.
 \item For uses of the \texttt{WorkStream} interface:
   \cite{TKB16}.
\end{itemize}

\marginpar{All: update gmsh reference}
\dealii{} can interface with many other libraries:
\begin{multicols}{3}
\begin{itemize}
\item ADOL-C \cite{adol-c}
\item ARPACK \cite{arpack}
\item Assimp \cite{assimp}
\item BLAS and LAPACK
\item Gmsh \cite{...}
\item GSL \cite{gsl2016}
\item HDF5 \cite{hdf5}
\item METIS \cite{karypis1998fast}
\item MUMPS \cite{ADE00,MUMPS:1,MUMPS:2,mumps-web-page}
\item muparser \cite{muparser-web-page}
\item nanoflann \cite{nanoflann}
\item NetCDF \cite{rew1990netcdf}
\item OpenCASCADE \cite{opencascade-web-page}
\item p4est \cite{p4est}
\item PETSc \cite{petsc-user-ref,petsc-web-page}
\item ROL \cite{ridzal2014rapid}
\item ScaLAPACK
\item SLEPc \cite{Hernandez:2005:SSF}
\item SUNDIALS \cite{sundials}
\item TBB \cite{Rei07}
\item Trilinos \cite{trilinos,trilinos-web-page}
\item UMFPACK \cite{umfpack}
\end{itemize}
\end{multicols}
Please consider citing the appropriate references if you use interfaces to these
libraries.

Older releases of \dealii{} can be cited as
\cite{dealII80,dealII81,dealII82,dealII83,dealII84,dealII85}.

@@

\dealii{} is a world-wide project with dozens of contributors around the
globe. Other than the authors of this paper, the following people contributed code to
this release:\\
%
% get this from the changes/*/* files using the command listed in the
% release-tasks paper and remove the authors of this paper
%
% Updated 2018-04-19
Julian Andrej,
Rajat Arora,
Lucas Campos,
Praveen Chandrashekar,
Jie Cheng,
Emma Cinatl,
Conrad Clevenger,
Ester Comellas,
Sambit Das,
Giovanni Di Ilio,
Nivesh Dommaraju,
Marc Fehling,
Niklas Fehn,
Menno Fraters,
Anian Fuchs,
Daniel Garcia-Sanchez,
Nicola Giuliani,
Anne Glerum,
Christoph Goering,
Alexander Grayver,
Samuel Imfeld,
Daniel Jodlbauer,
Guido Kanschat,
Vishal Kenchan,
Andreas Kergassner,
Eldar Khattatov,
Ingo Kligge,
Uwe Köcher,
Joachim Kopp,
Ross Kynch,
Konstantin Ladutenko,
Tulio Ligneul,
Karl Ljungkvist,
Santiago Ospina,
Alexey Ozeritsky,
Dirk Peschka,
Simon Puchert,
E. G. Puckett,
Lei Qiao,
Ce Qin,
Jonathan Robey,
Alberto Sartori,
Daniel Shapero,
Ben Shields,
Simon Sticko,
Oliver Sutton,
Zhuoran Wang,
Xiaoyu Wei,
Michał Wichrowski,
Julius Witte,
Feimi Yu,
Weixiong Zheng.

@@

\dealii{} and its developers are financially supported through a
variety of funding sources:
\marginpar{all: update your funding sources}

@@

The Interdisciplinary Center for Scientific Computing (IWR) at Heidelberg
University has provided hosting services for the \dealii{} web page.

>>>>

\usepackage{fontenc}
\usepackage{graphicx}
\usepackage{xspace}

@@

\newcommand{\specialword}[1]{\texttt{#1}}
\newcommand{\dealii}{{\specialword{deal.II}}\xspace}
\newcommand{\pfrst}{{\specialword{p4est}}\xspace}
\newcommand{\trilinos}{{\specialword{Trilinos}}\xspace}
\newcommand{\aspect}{\specialword{Aspect}\xspace}
\newcommand{\petsc}{\specialword{PETSc}\xspace}
\newcommand{\cmake}{{\specialword{CMake}}\xspace}
\newcommand{\autoconf}{{\specialword{autoconf}}\xspace}

@@

\title{The \dealii Library, Version 9.0}

@@

\begin{abstract}
  This paper provides an overview of the new features of the finite element
  library \dealii version 9.0.
\end{abstract}

@@

\dealii version 9.0.0 was released April X, 2018.
\marginpar{Update the date}
This paper provides an
overview of the new features of this major release and serves as a citable
reference for the \dealii software library version 9.0. \dealii is an
object-oriented finite element library used around the world in the
development of finite element solvers. It is available for free under the
GNU Lesser General Public License (LGPL) from the \dealii homepage at
\url{http://www.dealii.org/}.

The major changes of this release are:
\begin{itemize}
\item Improved support for curved geometries;
\item Support for particle-in-cell methods;
\item Dedicated support for automatic and symbolic differentiation;
\item Interfaces to more external libraries and programs;
\item C++11 is now both required and used;
\item Support for GPU computations;
\end{itemize}
These will all be discussed in more detail in the
following section. In addition, this release contains the following changes:
\begin{itemize}
\item \dealii has made extensive use of both the Clang-Tidy \cite{clang-tidy}
  and Coverity Scan \cite{coverity} static analysis tools for detecting bugs
  and other issues in the code. For example, around 260 issues were detected and
  fixed using the latter tool.
      %% > Is there somewhere we can cite for either of these tools? It would
      %% > be nice to recognize the department of homeland security for offering
      %% The contract with homeland security was completed in 2009.

@@

\item A new python tutorial program tutorial-1; as well as
  updates to step-37. In addition, the separate code
  gallery of \dealii has gained a number of new entries.

\item Improved support for user parameters; a new \texttt{ParameterAcceptor}
  class has been added to the library. The class is intended to be used as a
  base for any class that wants to handle parameters using the
  \texttt{ParameterHandler} class. If you derive all your classes from
  \texttt{ParameterAcceptor}, and declare your parameters either with
  \texttt{parse/declare\_parameters} methods or via the
  \texttt{ParameterAcceptor::add\_parameter} method, then the declaring and
  parsing of your parameter files will be automatically managed by the
  \texttt{ParameterAcceptor::initialize} function, greatly simplifying
  the declaration and parsing of parameters in user codes.

\item New caching mechanism for expensive grid computations; We introduced a new
  class, \texttt{GridTools::Cache}, that caches computationally intensive
  information about a \texttt{Triangulation}.

This class allows the user to query some of the data structures constructed
using functions in the GridTools namespace which are computed only once, and
then cached inside this class for faster access whenever the triangulation has
not changed. The Cache is marked for update by the \texttt{Triangulation} itself
using signals.

Some of the methods in \texttt{GridTools} already use the \texttt{Cache} to
speed up repeated calls to the same expensive methods.
\item A new \texttt{MeshWorker::mesh\_loop} function has been added that
  performs the same tasks of the \texttt{MeshWorker::loop} function without
  forcing the users to adhere to a specific interface.

\end{itemize}
Beyond these changes, the changelog lists more than 330 other features and bugfixes.

@@

This release of \dealii contains a number of large and significant changes
that will be discussed in the following sections. It of course also contains a
vast number of smaller changes and added functionality; the details of these
can be found
\href{https://www.dealii.org/developer/doxygen/deal.II/changes_between_8_5_and_9_0.html}{
in the file that lists all changes for this release}, see \cite{changes90}.
(The file is also linked to from the web site of each release as well as
the release announcement.)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Improved support for curved geometries}
\label{sec:manifolds}

@@

Every function in the GridGenerator namespace now attaches a default manifold to
the curved parts of the domain, and sets reasonable defaults for manifold
indicators both in the domain and on the boundary, where appropriate. Manifold
classes are no longer sensible to settings made on boundary indicators, and are
only affected by manifold indicators, decoupling completely the concept of
boundary indicators and manifold indicator, leaving the former for boundary
conditions, and the latter for the geometrical description of the domain.

All functions that allowed to query for \texttt{Boundary} objects (deprecated
since version 8.9 of the \dealii library) have been removed, and have been
replaced by the equivalent methods that query \texttt{Manifold} classes.

The manifold smoothing algorithms applied in the Triangulation class and
MappingQGeneric have been changed from the old Laplace-style smoothing to a
transfinite interpolation that linearly blends between the descriptions on the
faces around a cell. The old transformation introduced boundary layers inside
cells that prevented convergence rates from exceeding \(3.5\) in the global
\(L^2\) errors on typical settings. This change also considerably improves mesh
quality on settings where curved descriptions are only applied to the boundary
rather than the whole volume.

A new manifold class \texttt{TransfiniteInterpolationManifold} implementing an
interpolation from a curved boundary description to the interior has been added.
This class enables high-order convergence rates of more than three in the power
of the mesh size for situations where a curved manifold can only be prescribed
to the boundary but not in a whole volume.

@@

\marginpar{Rene: Please edit as appropriate}
While \dealii is a package intended to solve problems with the
finite element method -- i.e., using continuous or discontinuous
\textit{fields} --, is often convenient in fluid dynamics problems to
couple the continuum description of phenomena with particles. These
particles, advected along with the numerical approximation of the flow
field, are then either used to visualize properties of the flow, or to
advect material properties such as the viscosity of inhomogenous
mixtures of fluids. If each particle is associated with the cells
of a mesh, these methods are often referred to as particle-in-cell (PIC).

\dealii now has a dedicated particles module. The module provides a base class
\texttt{Particle} that represents a particle with position, an ID number and
a variable number of properties. They are jointly represented by a \texttt{ParticleHandler}
class that manages the storage and handling of all particles. In
parallel simulations, this class also
distributes the particles among the subdomains of the parallel process
and supports efficient data transfer during mesh refinement and
checkpoint/restart phases.

A much more detailed view of the underlying algorithms can be found in
\cite{GLHPB18}. A longer report is at \cite{GHPB16}. The
implementation here originated in the \aspect code, see \cite{KHB12,HDGB17}.

@@

\dealii has had a tutorial program (step-33) since 2007 that
demonstrates this technique based on the Trilinos Sacado package, but
the functionality was not available pervasively throughout
deal.II. This has changed with release 9.0 where support for
differentiation has been collected in one module that provides both
automatic and symbolic differentiation. Currently, this module relies
on either the ADOL-C and Sacado libraries, through a unified
interface. In practice, this support means that ADOL-C and Sacado data
types can be used in the \texttt{FEValues}, \texttt{FEValuesViews},
and related classes that are generally used to assemble linear systems
and right hand sides.

@@

\dealii has always tried to leverage high-quality implementations of
algorithms available through other open source software, rather than
re-implementing their functionality. (A list of interfaces to other
packages is given in Section~\ref{sec:cite}.) As part of the current
release, we have written several new interfaces as discussed in the following.

@@

  \dealii now has interfaces to SUNDIALS's ARKode, IDA, and KINSOL sub-packages.
  ARKode is a solver library that provides adaptive-step time
  integration. IDA is a package for the solution of differential-algebraic
  equations systems in the form $F(t,y,y')=0$. KINSOL is solver for nonlinear
  algebraic systems.

@@

\paragraph*{nanoflann, a library for building and querying
  $k$-d trees of datasets.} Operations such as finding the vertex
or cell closest
to a given evaluation point occur frequently in many applications that
use unstructured meshes. While the naive algorithm is linear in the
number of vertices or cells, many such operations can be made
significantly faster by building a $k$-d tree data structure that
recursively subdivides a $k$ dimensional space. The nanoflann library
provides such a data structure and allows querying it, either for
closest points (e.g., when finding the closest vertex) or for
searching the points that fall within a radius of a target point. This
functionality is now available via \dealii interfaces.

\paragraph*{Assimp, the Open Asset Import Library.}
  Assimp can be used to read about 40 different 3D graphics formats. A
  subset of these formats can be now be read from within \dealii to
  generate two-dimensional meshes, possibly embedded in a
  three-dimensional space.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \subsection{Use of C++11}
  \label{sec:cxx11}
\dealii first offered support for a subset of C++11 features in
version 6.2, released in 2009. The current release is the first to
\emph{require} a C++11 compiler.

Many parts of the code base have been rewritten to both support and
use the new features of C++11. In particular, \dealii now makes
extensive use of move semantics as well as range-based \texttt{for}
loops with \texttt{auto} type deduction of iterator variables. We have
also largely replaced \texttt{push\_back()} by
\texttt{emplace\_back()} when adding elements to collections more
efficiently.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{MeshWorker::mesh\_loop}

\marginpar{do we want this? LH}

@@

\dealii has a separate ``code gallery'' that
consists of programs shared by users as examples of what can be
done with \dealii. While not part of the release process, it is nonetheless
worth mentioning that the set of new programs since the last release covers
the following topics:
\begin{itemize}
\item The multipoint flux mixed finite element method (MFMFE) applied
  to the Darcy problem of porous media flow;
\item A linearized active skeletal muscle model with application to
  the simulation concentric contraction of the human biceps brachii;
\item A parallel implementation of the Local Discontinuous Galerkin
  (LDG) method applied to the Poisson equation.
\end{itemize}
With these additions, the code gallery now contains 10 different applications.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{How to cite \dealii}\label{sec:cite}

In order to justify the work the developers of \dealii put into this
software, we ask that papers using the library reference one of the
\dealii papers. This helps us justify the effort we put into it.

There are various ways to reference \dealii. To acknowledge the use of
the current version of the library, \textbf{please reference the present
document}. For up to date information and bibtex snippets for this document
see:
\begin{center}
 \url{https://www.dealii.org/publications.html}
\end{center}

The original \texttt{\dealii} paper containing an overview of its
architecture is \cite{BangerthHartmannKanschat2007}. If you rely on
specific features of the library, please consider citing any of the
following:
\begin{itemize}
 \item For geometric multigrid: \cite{Kanschat2004,JanssenKanschat2011};
 \item For distributed parallel computing: \cite{BangerthBursteddeHeisterKronbichler11};
 \item For $hp$ adaptivity: \cite{BangerthKayserHerold2007};
  \item For partition-of-unity (PUM) and enrichment methods of the
    finite element space: \cite{Davydov2016};
 \item For matrix-free and fast assembly techniques:
   \cite{KronbichlerKormann2012};
 \item For computations on lower-dimensional manifolds:
   \cite{DeSimoneHeltaiManigrasso2009};
 \item For integration with CAD files and tools:
   \cite{HeltaiMola2015};
 \item For \texttt{LinearOperator} and \texttt{PackagedOperation} facilities:
   \cite{MaierBardelloniHeltai-2016-a,MaierBardelloniHeltai-2016-b}.
 \item For uses of the \texttt{WorkStream} interface:
   \cite{TKB16}.
\end{itemize}

\marginpar{All: update gmsh reference}
\dealii can interface with many other libraries:
\begin{multicols}{3}
\begin{itemize}
\item ADOL-C \cite{adol-c}
\item ARPACK \cite{arpack}
\item Assimp \cite{assimp}
\item BLAS and LAPACK
\item Gmsh \cite{...}
\item GSL \cite{gsl2016}
\item HDF5 \cite{hdf5}
\item METIS \cite{karypis1998fast}
\item MUMPS \cite{ADE00,MUMPS:1,MUMPS:2,mumps-web-page}
\item muparser \cite{muparser-web-page}
\item nanoflann \cite{nanoflann}
\item NetCDF \cite{rew1990netcdf}
\item OpenCASCADE \cite{opencascade-web-page}
\item p4est \cite{p4est}
\item PETSc \cite{petsc-user-ref,petsc-web-page}
\item ROL \cite{ridzal2014rapid}
\item ScaLAPACK
\item SLEPc \cite{Hernandez:2005:SSF}
\item SUNDIALS \cite{sundials}
\item TBB \cite{Rei07}
\item Trilinos \cite{trilinos,trilinos-web-page}
\item UMFPACK \cite{umfpack}
\end{itemize}
\end{multicols}
Please consider citing the appropriate references if you use interfaces to these
libraries.

Older releases of \dealii can be cited as
\cite{dealII80,dealII81,dealII82,dealII83,dealII84,dealII85}.

@@

\dealii is a world-wide project with dozens of contributors around the
globe. Other than the authors of this paper, the following people contributed code to
this release:\\
%
% get this from the changes/*/* files using the command listed in the
% release-tasks paper and remove the authors of this paper
%
% Updated 2018-04-19
Julian Andrej,
Rajat Arora,
Lucas Campos,
Praveen Chandrashekar,
Jie Cheng,
Emma Cinatl,
Conrad Clevenger,
Ester Comellas,
Sambit Das,
Giovanni Di Ilio,
Nivesh Dommaraju,
Marc Fehling,
Niklas Fehn,
Menno Fraters,
Anian Fuchs,
Daniel Garcia-Sanchez,
Nicola Giuliani,
Anne Glerum,
Christoph Goering,
Alexander Grayver,
Samuel Imfeld,
Daniel Jodlbauer,
Guido Kanschat,
Vishal Kenchan,
Andreas Kergassner,
Eldar Khattatov,
Ingo Kligge,
Uwe Köcher,
Joachim Kopp,
Ross Kynch,
Konstantin Ladutenko,
Tulio Ligneul,
Karl Ljungkvist,
Santiago Ospina,
Alexey Ozeritsky,
Dirk Peschka,
Simon Puchert,
E. G. Puckett,
Lei Qiao,
Ce Qin,
Jonathan Robey,
Alberto Sartori,
Daniel Shapero,
Ben Shields,
Simon Sticko,
Oliver Sutton,
Zhuoran Wang,
Xiaoyu Wei,
Michał Wichrowski,
Julius Witte,
Feimi Yu,
Weixiong Zheng.

@@

\dealii and its developers are financially supported through a
variety of funding sources:
\marginpar{all: update your funding sources}

@@

The Interdisciplinary Center for Scientific Computing (IWR) at Heidelberg
University has provided hosting services for the \dealii web page.
